---
title: "Mid-Term Decision"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, message=FALSE, warning=FALSE}
pacman::p_load(readr, haven, readxl, downloader, tidyverse, ggbeeswarm, mosaic, stringr, pander, DT, ggplot2, alr3, foreign, measurements)

```

```{r Load Data, message=FALSE, warning=FALSE}
# NA = 0 in the gradebook
# Can make quantitative data into T/F statements
Midterm <- read.csv("../../Data/MidtermVsFinalGrades2.csv")

#View(Midterm)
```

```{r Data Wrangling, message=FALSE, warning=FALSE}
########## Cleaned data and created columns with people around me ####################
########## Found mutate_all here (https://stackoverflow.com/questions/19379081/how-to-replace-na-values-in-a-table-for-selected-columns) ######################

Midterm2 <- Midterm %>% 
  select(-"Gender") %>%
  mutate_all(~replace_na(., 0)) %>% 
  mutate(Analyses86 = ifelse(AnalysesTheory > 86, 1, 0),
#         AssesActual66 = ifelse(AssessmentQuizActual > 66, 1, 0),
#         Mid76 = ifelse(Midterm == 76, 1, 0),
         Mid72_80 = ifelse(Midterm >= 72 & Midterm < 80, 1, 0),
         Overall = round(Midterm*.06 + FinalExam*.14 + AssessmentQuizCompletion*.05 + AssessmentQuizActual*0 + AnalysesTheory*.3 + PeerReviews*.1 + ClassActivities*.1 + SkillQuizzes*.1 + 100*.15, digits = 2),
         Overall2.0 = round(((Midterm + AssessmentQuizCompletion + AssessmentQuizActual + AnalysesTheory + PeerReviews + ClassActivities + SkillQuizzes)/7), digits = 2)) %>% #View()
  mutate(Grade = as.factor(case_when(
    Overall >= 93 ~ "A",
    Overall >= 90 & Overall < 93 ~ "A-",
    Overall >= 87 & Overall < 90 ~ "B+",
    Overall >= 83 & Overall < 87 ~ "B",
    Overall >= 80 & Overall < 83 ~ "B-",
    Overall >= 77 & Overall < 80 ~ "C+",
    Overall >= 73 & Overall < 77 ~ "C",
    Overall >= 70 & Overall < 73 ~ "C-",
    Overall >= 67 & Overall < 70 ~ "D+",
    Overall >= 63 & Overall < 67 ~ "D",
    Overall >= 60 & Overall < 63 ~ "D-",
    Overall > 0 & Overall < 60 ~ "F")),
         Grade.group = as.factor(case_when(
    Overall >= 90 ~ "A",
    Overall >= 80 & Overall < 90 ~ "B",
    Overall >= 70 & Overall < 80 ~ "C",
    Overall >= 60 & Overall < 70 ~ "D",
    Overall > 0 & Overall < 60 ~ "F")))

#View(Midterm2)

########### Removed all the people who didn't take the Midterm #######################

Midterm_0_removed <- Midterm2[c(-3, -8, -26, -38, -53, -42), ] 
# 3 0% Midterm. Didn't take midterm. Everything else seems okay.
# 8 0% Midterm. Seems to have given a half-hearted effort.
# 26 0% Midterm. 67% on Theories and 52% on Final.
# 38 0% everything. Withdrawn?
# 42 Removed as person appears to have given up after the midterm?
# 53 0% on Mid/Final. Half-hearted effort rest of semester.


# Need to replace NA with zeros.
# View(Midterm2)
# View(Midterm_0_removed)
```

----

**Comments to Critiquers**

1. How do I reorder the levels of the `Grade` and `Grade.group` columns since there are +, -, and the letter e is missing?
2. Don't I want to base this entire analysis off of my midterm score? My `Overall` column seems to be a better option to use than `Midterm`, but I won't know my `Overall` score til the end of the semester. . . What do I do?
3. Do I remove people that are messing with the data?

**Corrections I made**

1. I added my Midterm score to the Summary section. (Saunders)
2. 

----

## Background

I received a 76 out of 100 on my midterm. I get to determine whether I will keep or drop my mid-term score as it makes up 30% of the 20% that my Midterm and Final scores count toward my final grade.  

## What You'd Like {.tabset .tabset-pills .tabset-fade}

### Summary {.tabset}

```{r eval=FALSE}
Mid10.a.lm <- lm(FinalExam ~ Midterm + ClassActivities, data = Midterm2)
predict(Mid10.a.lm, data.frame(Midterm = 76, ClassActivities = 100), interval = "prediction")
```

I am going to keep my midterm score of 76. My final exam score is projected to be a 72 with the ability of me scoring between 39 and 100.

Just for kicks here is the prediction from Attempt 8:

```{r}
mid3 <- Midterm2[c(-38, -2), ]

Mid8.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + SkillQuizzes + Overall, data = mid3)
pander(summary(Mid8.lm))

predict(Mid8.lm, data.frame(Midterm = 76, AnalysesTheory = 93, SkillQuizzes = 100, Overall = 90), interval = "prediction")
```

With a 76 for my final, 14 out of 15 on my Analyses and Theories, get 100% on my Skill Quizzes, and hypothetically earn that A- I'm aiming for, I will receive a 57 on the Final. When playing around with the prediction function, I think it's because there's an expectation from that `Overall` column that if I did well with my Analyses and SkillQuizzes I should have received a higher `Final` score than I'm projecting for myself. I actually think it can be put this way: If I received a 90 in the course then it is expected that I should have received a Final exam score of the following if I completed everything in the course 100%:

```{r}
predict(Mid8.lm, data.frame(Midterm = 76, AnalysesTheory = 100, SkillQuizzes = 100, Overall = 90), interval = "prediction")

predict(Mid8.lm, data.frame(Midterm = 76, AnalysesTheory = 80, SkillQuizzes = 100, Overall = 90), interval = "prediction")
```


### Analysis {.tabset .tabset-pills .tabset-fade}

#### Attempt 1 {.tabset}

```{r Pairs, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
pairs(Midterm2, panel = panel.smooth, pch = 16, cex = 0.8, col = "black")
pairs(Midterm_0_removed, panel = panel.smooth, pch = 16, cex = 0.8, col = "black")
```

Added columns `Analyses86`, `Mid72_80`, `Overall`, `Grade`, and `Grade.group`.

1. `Analyses86`: Those who got 13 out of 15.
2. `Mid72_80`: Those who scored similarly to me.
3. `Overall`: Overall course final grade. *Since no regression battleship column was available, everyone got a 100% to calculate this.*
4. `Grade`: Letter grade (11 levels)
5. `Grade.group`: Groupings of letter grades (5 levels)

I'm not sure if it's better to remove these observations or not. Here is what I see:

* 3 - 0% Midterm. Didn't take midterm. Everything else seems okay.
* 8 - 0% Midterm. Seems to have given a half-hearted effort.
* 26 - 0% Midterm. 67% on Theories and 52% on Final.
* 38 - 0% everything. Withdrawn?
* 42 - Person appears to have given up after the midterm?
* 53 - 0% on Mid/Final. Half-hearted effort rest of semester.

I believe `Midterm` and `Overall` will be the best ones to use. But can I use overall since I don't know what my final grade might be? Don't I want to base this off of my midterm score?

#### Attempt 2

```{r fig.width=10, fig.height=10}
Mid.lm <- lm(FinalExam ~ Midterm, data = Midterm2)
pander(summary(Mid.lm))

Mid.rem.lm <- lm(FinalExam ~ Midterm, data = Midterm_0_removed)
pander(summary(Mid.rem.lm))

pairs(cbind(R = Mid.lm$res, Fit = Mid.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")
#pairs(cbind(R = Mid.rem.lm$res, Fit = Mid.rem.lm$fit, Midterm_0_removed), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid.lm, which=1:2)
plot(Mid.lm$residuals, main = "Residuals vs Order")
#plot(Mid.lm, which = 1)
# qqPlot(Mid.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

# # Check Your Assumptions
# par(mfrow=c(1,3))
# plot(Mid.rem.lm, which=1:2)
# plot(Mid.rem.lm$residuals, type = "b", main = "Residuals vs Order")
# #plot(Mid.rem.lm, which = 1)
# # qqPlot(Mid.rem.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Mid.lm has `Midterm` with 7.86e-05. $R^2$ is 0.1974.

Mid.rem.lm has `Midterm` with 0.000814, but the y is not significant. I think I'll keep the zeros in then.

Assumptions: Surprisingly it appears that variance is okay, while linearity relation between x and y is not normal. The error terms are not normally distributed. I'm also not sure, but there may be a pattern to the error terms? Observation 42 was one that I thought to remove. Not sure if this is okay or not.

`AssessmentQuizActual`, `AnalysesTheory`, `ClassActivities`, `SkillQuizzes`, `Analyses86`, and `Overall` seem to be some of the biggest players. Of these I think that `AnalysesTheory` and `Overall` do the best in explaining things. Before I wrote things out though I did a lm() with `AssessmentQuizActual` and `AnalysesTheory`. 

#### Attempt 3

```{r fig.width=10, fig.height=10}
Mid3.lm <- lm(FinalExam ~ Midterm + AssessmentQuizActual + AnalysesTheory, data = Midterm2)
pander(summary(Mid3.lm))

pairs(cbind(R = Mid3.lm$res, Fit = Mid3.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid3.lm, which=1:2)
plot(Mid3.lm$residuals, main = "Residuals vs Order")
#plot(Mid3.lm, which = 1)
# qqPlot(Mid3.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Why did `FinalExam` lose its significance? Does this matter? Everything else shows significance. $R^2$ is 0.4101.

Assumptions: Observation 50 is causing problems. Variance again seems normal. Linearity relation between x and y is not okay. Error terms are not distributed normally. Potential pattern again.

Pairs: I was thinking of adding `Overall` to this, but the more I think about it, I think I shouldn't be using this column and my other other columns I created. I'll need to ask Brother Saunders. I am going to try look at all of them and their interactions.

#### Attempt 4

```{r eval=FALSE, fig.width=10, fig.height=10}
Mid4.lm <- lm(FinalExam ~ Midterm * AssessmentQuizActual * AnalysesTheory * ClassActivities * SkillQuizzes * Analyses86 * Overall, data = Midterm2)
pander(summary(Mid4.lm))

pairs(cbind(R = Mid4.lm$res, Fit = Mid4.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid4.lm, which=1:2)
plot(Mid4.lm$residuals, type = "b", main = "Residuals vs Order")
#plot(Mid4.lm, which = 1)
# qqPlot(Mid4.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Well that didn't work. But I have an $R^2$ of 1! Haven't seen that before. Let's look at Attempt 2 again. I'm going to try `AnalysesTheory` and `Skillquizzes`.

#### Attempt 5

```{r fig.width=10, fig.height=10}
Mid5.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + SkillQuizzes, data = Midterm2)
pander(summary(Mid5.lm))

pairs(cbind(R = Mid5.lm$res, Fit = Mid5.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid5.lm, which=1:2)
plot(Mid5.lm$residuals, main = "Residuals vs Order")
#plot(Mid5.lm, which = 1)
# qqPlot(Mid5.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

I don't know what I'm doing at this point. What happens if I add `Overall`?

#### Attempt 6

```{r fig.width=10, fig.height=10}
Mid6.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + SkillQuizzes + Overall, data = Midterm2)
pander(summary(Mid6.lm))

pairs(cbind(R = Mid6.lm$res, Fit = Mid6.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid6.lm, which=1:2)
plot(Mid6.lm$residuals, main = "Residuals vs Order")
#plot(Mid6.lm, which = 1)
# qqPlot(Mid6.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Assumptions: Observation 38 seems to be problamatic. I'm going to remove it and see what happens.

#### Attempt 7

```{r fig.width=10, fig.height=10}
mid2<- Midterm2[-38, ]

Mid7.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + SkillQuizzes + Overall, data = mid2)
pander(summary(Mid7.lm))

pairs(cbind(R = Mid7.lm$res, Fit = Mid7.lm$fit, mid2), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid7.lm, which=1:2)
plot(Mid7.lm$residuals, main = "Residuals vs Order")
#plot(Mid7.lm, which = 1)
# qqPlot(Mid7.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

I first removed observation 38, but then observation 2 looks bad too. SO let's see.

#### Attempt 8

```{r fig.width=10, fig.height=10}
mid3 <- Midterm2[c(-38, -2), ]

Mid8.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + SkillQuizzes + Overall, data = mid3)
pander(summary(Mid8.lm))

pairs(cbind(R = Mid8.lm$res, Fit = Mid8.lm$fit, mid3), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid8.lm, which=1:2)
plot(Mid8.lm$residuals, main = "Residuals vs Order")
#plot(Mid8.lm, which = 1)
# qqPlot(Mid8.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Everything looks **really** good. Did I mess something up here or is this the final model? Even $R^2$ is at 0.8569. So I'm going to roll with this. Here's the mathmatical model:

$$
  \underbrace{\hat{Y_i}}_\text{Average Final Score} = -165.1 + -0.2423 \underbrace{X_{1i}}_\text{Midterm} + -2.087 \underbrace{X_{2i}}_\text{AnalysesTheory} + -0.4726 \underbrace{X_{3i}}_\text{SkillQuizzes} + 5.358 \underbrace{X_{4i}}_\text{Overall}
$$

#### Attempt 9

I realized when writing out that model that I can't control what my overall amount will be in order to predict my `Final` exam score. So I am going to look at Attempt 2 again, see what looks good, decide what I can control, and do that model.

It looks like the AnalysesTheory and ClassActivities all look promising. I am going to try that first.

```{r fig.width=10, fig.height=10}
Mid9.lm <- lm(FinalExam ~ Midterm + AnalysesTheory + ClassActivities, data = Midterm2)
pander(summary(Mid9.lm))

pairs(cbind(R = Mid9.lm$res, Fit = Mid9.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

#boxCox(Mid9.lm)

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid9.lm, which=1:2)
plot(Mid9.lm$residuals, main = "Residuals vs Order")
#plot(Mid9.lm, which = 1)
# qqPlot(Mid9.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)
```

Do I need to do a transformation on `AnalysesTheory`? Not sure why this isn't working. Also, the boxCox isn't working because y is negative? Why? Do I just need to remove `AnalysesTheory`?

Assumptions: Does 53 need to be removed?

Pair: Do I need to combine `AssessmentQuizActual` and `ClassActivities`?

#### Attempt 10

```{r fig.width=10, fig.height=10}
Mid10.lm <- lm(FinalExam ~ Midterm + AssessmentQuizActual + ClassActivities, data = Midterm2)
pander(summary(Mid10.lm))

pairs(cbind(R = Mid10.lm$res, Fit = Mid10.lm$fit, Midterm2), panel = panel.smooth, pch=16, cex=0.7, col="black")

#boxCox(Mid10.lm)

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid10.lm, which=1:2)
plot(Mid10.lm$residuals, main = "Residuals vs Order")
#plot(Mid10.lm, which = 1)
# qqPlot(Mid10.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

plot(FinalExam ~ Midterm, data=Midterm2, pch=16, col="darkgray")
lines(lowess(Midterm2$Midterm, Midterm2$FinalExam), col="firebrick")
```

I think I'm going to remove the zero people. They're throwing things off. 

```{r fig.width=10, fig.height=10}
Mid10.a.lm <- lm(FinalExam ~ Midterm + ClassActivities, data = Midterm2)
pander(summary(Mid10.a.lm))
```

I'm going to go with this model for something to show.

#### Attempt 11

```{r fig.width=10, fig.height=10}
Mid.0 <- Midterm_0_removed %>% 
  select(c(-Analyses86, -Mid72_80, -Overall, -Grade, -Grade.group)) %>% view()

Mid.rem.lm <- lm(FinalExam ~ Midterm, data = Mid.0)
pander(summary(Mid.rem.lm))

pairs(cbind(R = Mid.rem.lm$res, Fit = Mid.rem.lm$fit, Mid.0), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid.rem.lm, which=1:2)
plot(Mid.rem.lm$residuals, main = "Residuals vs Order")
#plot(Mid.rem.lm, which = 1)
# qqPlot(Mid.rem.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

```

This seems to be much better. I'm going to run this on `AssessmentQuizActual` and `AnalysesTheory`. 

#### Attempt 12

```{r fig.width=10, fig.height=10}
Mid.12.lm <- lm(FinalExam ~ Midterm + AssessmentQuizActual + AnalysesTheory, data = Mid.0)
pander(summary(Mid.12.lm))

pairs(cbind(R = Mid.12.lm$res, Fit = Mid.12.lm$fit, Mid.0), panel = panel.smooth, pch=16, cex=0.7, col="black")

# Check Your Assumptions
par(mfrow=c(1,3))
plot(Mid.12.lm, which=1:2)
plot(Mid.12.lm$residuals, main = "Residuals vs Order")
#plot(Mid.12.lm, which = 1)
# qqPlot(Mid.12.lm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

```

I give up. I don't know what to do at this point.

### Data

This data was provided to me by Brother Saunders. As a reminder, by downloading and using this data I have agreed to:

1. Keep the raw data confidential (though I can share my analysis of the data with others)
2. Not try to retrace any of the data to the original individuals in order to protect their rights and privacy.

```{r Datatable, message=FALSE, warning=FALSE}
datatable(Midterm)
```

