---
title: "Assessment Quiz: Residuals, Sums of Squares, and R-squared"
author: "Devin Bastian"
date: "1/21/2020"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

## Assessment Quiz: Residuals, Sums of Squares, and R-squared


### Question 1

A regression was performed for a sample of n = 5 data points.

The y-values of the regression are: 3.78, 6.08, 6.65, 9.25, and 9.92.

The residuals from the regression are: -0.266, 0.489, -0.486, 0.569, and -0.306.

What is the R-squared value for this regression?

```{r}
y <- c(3.78, 6.08, 6.65, 9.25, 9.92)

SSTO <- sum( (y - mean(y))^2 )

#SSTO = 24.83372

res <- c(-0.266, 0.489, -0.486, 0.569, -0.306)

SSE <- sum(res^2)

#SSE = 0.96347

R2 = 1 - SSE/SSTO

#R2 = 0.9612032
```

### Question 2

Open the mtcars data set in R.

This data can be used to show that the displacement of the engine (disp) is positively correlated with the weight of the vehicle (wt).

What is the R-squared value of this regression?

```{r}
mt.lm <- lm(disp ~ wt, data=mtcars)

summary(mt.lm)
```

### Question 3

Open the mtcars data set in R.

Perform a regression of mpg on weight of the vehicle (mpg ~ wt).

A certain statistics teacher at BYU-Idaho drives a 2001 Nissan Sentra that weighs approximately 2,700 lbs and currently gets only 21 mpg. Based on the regression you performed, how many mpg above or below average is this vehicle?

```{r}
mt.lm <- lm(mpg ~ wt, data=mtcars) #perform the regression

plot(mpg ~ wt, data=mtcars) #draw the regression (not needed, but nice)

abline(mt.lm) #add the regression line (not needed, but nice)

points(2.7, 21, pch=16, col="skyblue") #add the Nissan Sentra value (not needed, but nice)

predict(mt.lm, data.frame(wt=2.7)) #get predicted value for Nissan Sentra

1

22.85505

lines(c(2.7, 2.7), c(21, 22.85505), col="skyblue") #add residual line (not needed, but nice)

myresidual <- 21 - 22.85505 #calculate difference between Y and Y-hat

myresidual
# [1] -1.85505
```

