---
title: "Practice Exam 1"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
---

```{r Libraries, message=FALSE, warning=FALSE}
library(car)
library(tidyverse)
library(mosaic)
library(pander)
library(ggplot2)
library(readr)
library(stringr)
library(DT)
library(pander)
library(gridExtra)
library(MASS)
library(ResourceSelection)
```


---

#### Question 1

To get a p-value you need two things. What are they?
Stat and distribution

```{r}

```



#### Question 2

How many runners in the race were from North Carolina (NC), New Jersey (NJ), and Maryland (MD)?

```{r}
#View(TenMileRace)

TenMileRace2 <- filter(TenMileRace, state %in% c("NC", "NJ", "MD"))

TenMileRace2 %>%

   group_by(state) %>%

   summarise(Count = n())

```




#### Question 3

Perform an analysis that fits the following mathematical model to the cats data set.

$$
  \underbrace{Y_i}_\text{heart Weight} = \overbrace{\beta_0}^\text{y-int} + \overbrace{\beta_1}^\text{slope} \underbrace{X_i}_\text{Body Weight} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$

State the estimate of LaTeX: $\beta_1$ from your analysis. 

```{r}
#View(cats)
#?cats

# Perform the Test
cats.mylm <- lm(Hwt ~ Bwt, data = cats)
summary(cats.mylm)


# Check Assumptions
par(mfrow=c(1,3))
plot(cats.mylm, which=1:2)
plot(cats.mylm$residuals, main = "Residuals vs Order")
#plot(mylm, which = 1)
# qqPlot(mylm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

#Plot the Regression Line
#plot(Y ~ X, data=YourDataSet)
#abline(mylm)
```


#### Question 4

Which of the following options would be most appropriate for identifying any differences in the data depicted in this graph?

```{r}
#View(CES11)

boxplot(weight ~ education, data=CES11)

# Perform the ANOVA
CES.myaov <- aov(weight ~ education, data=CES11) 

# View the ANOVA Results
summary(CES.myaov)  

# Check ANOVA Assumptions
par(mfrow = c(1,2))
plot(CES.myaov, which=1:2)
#DATA IS SKEWED

pander(kruskal.test(weight ~ education, data=CES11))

```


#### Question 5

Based on the information in this data set, is there significant evidence (at the $\alpha$ = 0.05) that the presence of a speed limit reduces the daily numbers of traffic accidents?


```{r}
View(Traffic)
?Traffic

yes <- filter(Traffic, limit == "yes")

t.test(y ~ limit, data = Traffic, mu = 0, alternative = "two.sided", conf.level = 0.95)

#To perform a logistic regression:

traffic.glm <- glm(limit == "yes" ~ y, data = Traffic, family = binomial)
summary(traffic.glm)

Traffic %>%

   group_by(y) %>%

   summarise(avg = n(limit)) %>%
  pander()

#Goodness of fit test with many replicated x-values:
pchisq(232.72, 182, lower.tail = FALSE)

#Goodness of fit test with few or no replicated x-values
library(ResourcesSelection)
hoslem.test(traffic.glm$y, traffic.glm$fitted)

#To predict the probability that Yi=1 for a given x-value:
predict(YourObjectName, data.frame(ColumnName = DesiredValue), type = "response")

# Challenger Example
predict(chall.glm, data.frame(Temp=31), type='response')

#To predict by-hand
exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x))

#To find the estimated value of eβ1
exp(b1)
```


#### Question 6

Create a graphic in R that shows the distribution of Composition scores for painters that studied in the various schools of art. Perform an appropriate test of the data contained in your graphic that would allow for us to conclude whether or not the distribution of composition scores for painters from each of these schools can all be considered to be from the same population of scores.

```{r}
View(painters)

?painters


```



#### Question 7

Which gender is more likely to volunteer, men or women?

```{r}
View(Cowles)

cow <- filter(Cowles, volunteer == "yes")

pander(table(cow$sex, cow$volunteer), split.table=Inf)
x <- table(Cowles$sex, Cowles$volunteer)
x.chi <- chisq.test(x)
pander(x.chi)
pander((x.chi$expected), split.table=Inf)
pander((x.chi$expected > 5), split.table=Inf)
pander((x.chi$residuals), split.table=Inf)
```


#### Question 8

Perform a permutation one-way ANOVA test to see if the average number of words correctly recalled is the same for each condition of the study. Report the p-value of your test.

```{r}
View(Friendly)

set.seed(101)

# Again, we run the initial test and find the test statistic
myTest <- aov(weight ~ feed, data = chickwts)
observedTestStat <- summary(myTest)[[1]]$`F value`[1]

# For this permutation, we need to shake up the groups similar to the Independent Sample example
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(chickwts$weight)
  permutedTest <- aov(permutedData ~ feed, data = chickwts)
  permutedTestStats[i] <- summary(permutedTest)[[1]]$`F value`[1]
}

# The histogram of this distribution gives an interesting insight into the results
hist(permutedTestStats, col = "skyblue", xlim = c(0,16))
abline(v = observedTestStat, col = "red", lwd = 3)

# Here is the greater-than p-value:
sum(permutedTestStats >= observedTestStat)/N

# Here is the less-than p-value: not the one that we want
sum(permutedTestStats <= observedTestStat)/N

# And here is the two-sided p-value: the same used in the original analysis
2*sum(permutedTestStats >= observedTestStat)/N


```


#### Question 9 

Filter the storms data set down to a data set that contains just hurricanes from 2015 during the month of August. This should result in 14 rows of data with only  hurricanes "Danny" and "Fred" remaining in this filtered data set.

Report the p-value of your test.
```{r}
View(storms)
?storms
```



#### Question 10

Run logistic curve

```{r}
?mtcars
#To perform a logistic regression:

cars.glm <- glm(am == 1 ~ mpg, data = mtcars, family = binomial)
summary(cars.glm)

plot(am == 1 ~ mpg, data = mtcars, xlab="Average Temperature", ylab="Did $20 Cover The Bill?", main="Success or Failure of $20 Fitting the Electric Bill", col="chocolate1", pch=19)
curve(exp(-6.6035 + 0.3070*x)/(1 + exp(-6.6035 + 0.3070*x)), add = TRUE)

#Goodness of fit test with many replicated x-values:
pchisq(residual deviance, df for residual deviance, lower.tail = FALSE)

#Goodness of fit test with few or no replicated x-values
#library(ResourcesSelection)
hoslem.test(YourObjectName$y, YourObjectName$fitted)

#To predict the probability that Yi=1 for a given x-value:
predict(YourObjectName, data.frame(ColumnName = DesiredValue), type = "response")

# Challenger Example
predict(chall.glm, data.frame(Temp=31), type='response')

#To predict by-hand
exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x))

#To find the estimated value of eβ1
exp(b1)

```



#### Question 11

```{r}

```



#### Question 12


```{r}
View(Chile)

# Perform the Test
chile.mylm <- lm(population ~ income, data = Chile)
summary(chile.mylm)


# Check Assumptions
par(mfrow=c(1,3))
plot(chile.mylm, which=1:2)
plot(chile.mylm$residuals, main = "Residuals vs Order")
#plot(mylm, which = 1)
# qqPlot(mylm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

#Plot the Regression Line
plot(population ~ income, data = Chile)
abline(chile.mylm)
```



#### Question 13

```{r}
View(SaratogaHouses)
# Perform the Test
homes.mylm <- lm(price ~ heating + centralAir + heating + centralAir, data = SaratogaHouses)
pander(summary(homes.mylm))

# Check Your Assumptions
par(mfrow=c(1,3))
plot(mylm, which=1:2)
plot(mylm$residuals, type = "b", main = "Residuals vs Order")
#plot(mylm, which = 1)
# qqPlot(mylm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

# Plot Regression Line
predict(mylm, data.frame(X1=X1h, X2=X2h), type="response")
myPlot(NameOfYourDataset)
```



#### Question 14

```{r}

homes <- filter(SaratogaHouses, heating=="hot water/steam")
homes %>%

   group_by(centralAir) %>%

   summarise(Count = mean(price))
```


#### Question 15

```{r}
  
```



#### Question 16

```{r}

```



#### Question 17

```{r}

```


#### Question 18


```{r}
View(airquality)

mplot(airquality)

?plot

plot(Temp ~ Wind, 
data=airquality, 
xlab="Daily Wind Speed (mph)", ylab="Temperature", main="La Guardia Airport (May - Sep)", col="ivory3", pch=18)

gf_point(Temp ~ Wind, data = airquality, color = ~ Month) %>% 
gf_theme(legend.position = "right") %>% 
gf_labs(title = "", caption = "")
```



#### Question 19

compute the correlation between the length of a foot and the width of a foot for fourth grade students.

```{r}
?KidsFeet
# Perform the Test
feet.mylm <- lm(length ~ width, data = KidsFeet)
summary(feet.mylm)

feet2.mylm <- lm(width~length, data = KidsFeet)
summary(feet2.mylm)
# Check Assumptions
par(mfrow=c(1,3))
plot(mylm, which=1:2)
plot(mylm$residuals, type = "b", main = "Residuals vs Order")
#plot(mylm, which = 1)
# qqPlot(mylm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

#Plot the Regression Line
plot(Y ~ X, data=YourDataSet)
abline(mylm)
```



#### Question 20

```{r}


```





#### Question 21

```{r}
((340+315+180+100+50+28+10+10+390+310+170+70+20+10+2)-(340+315+180+100+50+28+10+10+390+310+170+70+20))/(340+315+180+100+50+28+10+10+390+310+170+70+20+10+2)
```


#### Question 22

```{r}
View(Prestige)
Prestige2 <- mutate(Prestige, type = as.factor(ifelse(type == "prof", "prof","other")))
View(Prestige2)

# Perform the Test
prestige.mylm <- lm(income ~ prestige + type + prestige:type, data = Prestige2)
pander(summary(prestige.mylm))

# Check Your Assumptions
par(mfrow=c(1,3))
plot(mylm, which=1:2)
plot(mylm$residuals, type = "b", main = "Residuals vs Order")
#plot(mylm, which = 1)
# qqPlot(mylm$residuals, main = "Q-Q Plot of Residuals", id = FALSE)

# Plot Regression Line
predict(mylm, data.frame(X1=X1h, X2=X2h), type="response")
myPlot(NameOfYourDataset)
```


#### Question 23

```{r}
View(KidsFeet)

boys.feet <- filter(KidsFeet, birthyear == 88 & sex == "B")
View(boys.feet)

wilcox.test(length ~ domhand, data = boys.feet, mu = 0, alternative = "two.sided", conf.level = 0.95)
```


#### Question 24

A recent `college graduate` just landed a new job and is trying to `find` the `quickest commute route` to work. A quick internet search lists `five different routes` they could take and shows the `average time` for each route (at the given moment). Curious, the student decides to check up on the "internet" to see how `their actual commuting times` compare to the `"internet's" predicted` times. Over the first week of their new job, they take each of the five different routes to work, `one route each day`, and record both the `internet's predicted` times and the `actual time` it took them to make the commute. They then perform an analysis to see how "reality" matches up with the internet's predictions.

```{r}

```


#### Question 25

```{r}
((9-5.035)^2)/5.035
chisq.test(8)
```

