---
title: "Practice Exam 1"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
---

```{r Libraries, message=FALSE, warning=FALSE}
library(car)
library(tidyverse)
library(mosaic)
library(pander)
library(ggplot2)
library(readr)
library(stringr)
library(DT)
library(pander)
library(gridExtra)
```

#### Question 1

```{r}
#Average of Wookiees, Gungans, or Kaminoans are taller.

#View(starwars)
starwars2 <- filter(starwars, species %in% c("Wookiee", "Gungan", "Kaminoan"))
View(starwars2)
starwars2 %>%
  group_by(species) %>% 
   summarise(avgheight= mean(height))

```

```{r}
# Correcting from Attempt 1
starwars2 <- filter(starwars, species %in% c("Wookiee", "Gungan", "Kaminoan"))
xyplot(height ~ species, data=starwars2, main="", type=c("p","a")) 
star.wars.aov <- aov(height ~ species, data=starwars2)
summary(star.wars.aov)
```



#### Question 2

<!--  H0:μ2 Lanes=μ4 Lanes

      Ha:μ2 Lanes≠μ4 Lanes -->


```{r}
#View(Highway1)
#?Highway1

#t.test(Y (numberic) ~ X (qualitative), data = YourData, mu = YourNull, alternative = YourAlternative, conf.level = 0.95)

Highway2 <- Highway1 %>% 
  mutate(lanes = as.factor(lane))

#View(Highway2)

Highway2 %>%
  group_by(lanes) %>% 
   summarise(`Accident Count` = n())

t.test(Highway2$rate[Highway2$lanes == 2], Highway2$rate[Highway2$lanes == 4], mu = 0, alternative = "two.sided", conf.level = 0.95)

par(mfrow=c(1,2))

qqPlot(Highway2$rate[Highway2$lanes == 4])

qqPlot(Highway2$rate[Highway2$lanes == 2])

#t.test(Highway1$rate(Highway1$as.factor(lane == 2)) ~ Highway1$rate(Highway1$as.factor(lane == 4)), mu = 0, alternative = "two.sided", conf.level = 0.95)


```




#### Question 3

```{r}

plot(height ~ age, data=Loblolly)

lolly.lm <- lm(height ~ age, data = Loblolly)
pander(summary(lolly.lm))


par(mfrow = c(1,3),
    split.graph = Inf)
plot(lolly.lm,
     which = 1:2)
plot(lolly.lm$residuals, 
     main = "Residuals vs Order")
```


#### Question 4

Which analysis would allow you to predict whether or not it will rain on a given day based on the low temperature of that day?  
Logistic Regression

#### Question 5

Why might a Wilcoxon Rank Sum Test not be appropriate for testing to see if the median width of a child's foot differs for boys and girls?

```{r}
#View(KidsFeet)
pander(table(KidsFeet$sex, KidsFeet$width))
```


#### Question 6

```{r}
#View(Highway1)

Highway1 %>% 
   group_by(htype) %>% 
   summarise(`Avg. Accident` = mean(rate))
```



#### Question 7

What's the appropiate p-value based upon on the graph?

#### Question 8

```{r}
View(KidsFeet)

KidsFeet %>% 
   group_by(name) %>% 
   summarise(`Avg. Width` = mean(width),
             `Count of Names` = n()) %>%
  pander()
```



#### Question 9 

```{r}
View(iris)


Sepal.lm <- lm(Sepal.Length ~ Sepal.Width + Species + Sepal.Width:Species, data=iris)
pander(summary(Sepal.lm))
mPlot(iris)
plot(Sepal.Length ~ Sepal.Width, data=iris)
abline(Sepal.lm,lty=1,lwd=1,col="firebrick")

ggplot(data = iris, aes(x = Sepal.Width, y = Sepal.Length)) + geom_point()  + aes(size = Species)  + stat_smooth(method = lm) + theme(legend.position = "right") + labs(title = "")


(2.639+0.9007)
(0.6905+0.1746)

```

#### Question 10

Business majors more likely to marry than three other majors?

```{r}

```

#### Question 11

```{r}
View(cars)
?cars

cars.lm <- lm(speed ~ dist, data=cars)
pander(summary(cars.lm))
predict(cars.lm, data.frame(dist = 130))
−17.58+3.93*30
−17.58+3.93*35
−17.58+3.93*25
−17.58+3.93*40

```


<!-- ##############################################################################################################################
                        Try Number 1: One and half hours and 32%
################################################################################################################################################-->


#### Question 12

HELP

Is the average height of Males different than that of the average height of Females?

```{r}
View(Galton)

# First run the initial test and gain the test statistic:
myTest <- t.test(height ~ sex, data = Galton, mu = 0)
observedTestStat <- myTest$statistic
observedTestStat

# Now we run the permutation for a distribution of test statistics
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(Galton$sex)
  permutedTest <- t.test(height ~ permutedData, data = Galton, mu = 0)
  permutedTestStats[i] <- permutedTest$statistic
}
permutedTestStats


# Now we create a histogram of that distribution
hist(permutedTestStats, col = "skyblue")
abline(v = observedTestStat, col = "red", lwd = 3)

#Greater-Than p-value: Not the correct one in this case
sum(permutedTestStats >= observedTestStat)/N

# Less-Than p-value: This is the corret one
sum(permutedTestStats <= observedTestStat)/N

# Two-Sided p-value
2*sum(permutedTestStats <= observedTestStat)/N
```

#### Question 13

```{r}
View(Highway2)
wilcox.test(Highway2$rate[Highway2$lanes == 2], Highway2$rate[Highway2$lanes == 4], mu = 0, alternative = "two.sided", conf.level = 0.95)
```



#### Question 14

```{r}
View(Highway1)
?Highway1
plot(rate ~ slim, data=Highway1, pch=16, xlab="", ylab="", main="")
```


#### Question 15

```{r}
View(singer)
?singer


kruskal.test(height ~ voice.part, data=singer)

singer %>% 
   group_by(voice.part) %>% 
   summarise(Middle = median(height),
             Avg = mean(height),
             `Count of Names` = n()) %>%
  pander()
pander(summary(singer))
  
```



#### Question 16

```{r}
View(cars)

plot(speed > 15 ~ dist, data=cars, ylab="Probability Speed > 15 mph", xlab="Stopping Distance (feet)")

cars.glm <- glm(speed > 15 ~ dist, data = cars, family = "binomial")
pander(summary(cars.glm))

predict(cars.glm, data.frame(dist=70), type = "response")
1-(exp(0.8142148))
```



#### Question 17

Which of the following is a nonparametric test?
Permutation/Randomization

#### Question 18

Perform a two-way ANOVA using the KidsFeet dataset in R that will allow you to test if the pattern shown in the following plot is real. Use the α=0.05 level and report the p-value of the test.

```{r}
xyplot(length ~ sex, data=KidsFeet, group=domhand, type=c("p","a"), auto.key=TRUE)

View(KidsFeet)
myaov <- aov(length ~ sex+domhand+sex:domhand, data=KidsFeet)
summary(myaov)


```



#### Question 19

Which set of statistics would be most meaningful to include in an analysis containing this histogram?

summary()

```{r}
hist(islands, xlab="Area in Thousands of Square Miles", main="Areas of the World's Major Landmasses")
```




#### Question 20

Which of the following plots can be created using the mtcars dataset in R?

```{r}
mPlot(mtcars)
?mtcars
View(mtcars)

ggplot(mtcars, aes(x = factor(am), fill = factor(cyl))) +
  geom_bar(position = "dodge", color = "black") +
  ggtitle("BYU-Idaho Faculty Applicant Amount By Referral Source") +
  labs(x="Referral Source", y="Count of Applicants") +
  scale_fill_discrete(name = "Department") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 12, hjust = 0.5, vjust=0.8))



singer %>% 
   group_by(am) %>% 
   summarise(Middle = median(height),
             Avg = mean(height),
             `Count of Names` = n()) %>%
  pander()

```





#### Question 21

What two things are needed to compute a p-value?

Test stat and distribution

#### Question 22

```{r}
exp(0.251)
```


#### Question 24

```{r}
((400+500+380+250+175+130+50)-2000)/2000
```


#### Question 25

```{r}
View(ToothGrowth)
Tooth.myaov <- aov(len ~ as.factor(dose)+supp+as.factor(dose):supp, data=ToothGrowth)
t.test(len ~ supp, data = ToothGrowth, mu = 0, alternative = "two.sided", conf.level = 0.95) 

```

