---
title: "High School Seniors t Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

<!-- Week 3 Anaylsis -->

```{r Library and Data, message=FALSE, warning=FALSE}
library(mosaic)
library(tidyverse)
library(pander)
library(ggplot2)
library(car)
library(DT)
library(readr)
library(readxl)
library(stringr)
library(dplyr)
library(expss)

HSS <- read_csv("../../Data/HighSchoolSeniors.csv")
#Remember: select "Session, Set Working Directory, To Source File Location", and then play this R-chunk into your console to read the HSS data into R. 
```

<!-- Instructions

1. Use the HSS dataset in R to come up with a question that can be answered with a t Test. 

Here is a link to the survey given to the students that participated in the study:
https://ww2.amstat.org/censusatschool/pdfs/C@SQuestionnaire.pdf
(Use the link to learn about what each column of data in the data set represents.)

2. Establish why your question is interesting and rephrase the question using a statistical "null" and "alternative" hypothesis. (Hint, see the "Math 325 Notebook -> R Help -> R-Markdown Hints" page of your textbook for instructions on how to write mathematical hypotheses.)

3. Clearly answer your question using statistics. Be sure to compute a p-value and provide supporting numerical and graphical summaries.

Note: you can create "groups" from the data in many ways. For example, suppose you wanted to create a variable that grouped people according to whether they thought recycling was more important than conserving water. You could do this by:

-->

----

<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

**Comments to Critiquers:**  

What I've done:

1. A data table was added to this analysis allowing you to now look at the original and modified data.
2. A link to the U.S. Census at School survey was added under the *Data* section.
3. `split.table = Inf` was added to my t.test summary.
4. `Outdoor_Activities_Hours ~ Allergies, data=HSSAllergies` was added to my QQ plots.
5. Wording was added to the Independent $t$-test summary and QQ-plot graphs.
6. Language about the QQ plots showing the data to be heavy-winged was updated to right-skewed.
7. Titles on summary table was updated.
8. The summary under the boxplot had Min, Q1, Q3, and Max added onto to it.
9. Conclusion paragraph was updated.

Questions I have:

* Why is there so much space after my Conclusion?
* What does `include=` in my R chunks do?

<!--End your comments to your critiquers before this line.-->

----

# Who Is Outside More?
<br>
I am comparing those with allegeries and how much time they spend outside. The null hypothesis is that people with and without allergies spend an equal amount of time outdoors. However, I propose that this will be incorrect that children with allergies will spend less time outside than their non-runny nose, non-teary eyed, and clear headed counterparts. In other words, those without allergies will spend more time outside than those without.

$$
  H_0: \underbrace{\mu_1}_\text{Without Allergies}  - \underbrace{\mu_2}_\text{With Allergies} = 0
$$
$$ 
  H_a: \underbrace{\mu_1}_\text{Without Allergies}  - \underbrace{\mu_2}_\text{With Allergies} > 0
$$

## Understanding The Data {.tabset .tabset-pills .tabset-fade}

### Summary 

```{r Boxplot, message=FALSE, warning=FALSE}
HSSAllergies <- filter(HSS, 
                       Allergies %in% c("Yes", "No") 
                       & Outdoor_Activities_Hours <= 168) 

#mPlot(HSSAllergies)

gf_boxplot(Outdoor_Activities_Hours ~ Allergies, data = HSSAllergies) %>% gf_labs(title = "Time People Spend Outdoors With and Without Allergies", x="Allergies", y="Hours Spent Outdoors")
```


```{r Summary Table, message=FALSE, warning=FALSE}

HSSAllergies <- filter(HSS, 
                       Allergies %in% c("Yes", "No") 
                       & Outdoor_Activities_Hours <= 168) 
#HSSAllergies2 <- select(HSS, c(Allergies))
pander(HSSAllergies %>%
  group_by(Allergies) %>%
  summarise(`Avg. Outdoor Time` = mean(Outdoor_Activities_Hours,
                                    na.rm="TRUE"),
            Min = min(Outdoor_Activities_Hours, na.rm="TRUE"), 
            Q1 = quantile(Outdoor_Activities_Hours, na.rm="TRUE", 0.25),
            `Median of Outdoor Time` = median(Outdoor_Activities_Hours, 
                                  na.rm="TRUE"),
            Q3 = quantile(Outdoor_Activities_Hours, na.rm="TRUE", 0.75),
            max = max(Outdoor_Activities_Hours, na.rm="TRUE"),
            `Sample Size` = n( )),
  split.table = Inf,
  caption = "Summary of Boxplot")
            
     
```

<br>

The average times reported by those with allergies surprisingly show that they spend more time outdoors, which is a curious observation. It leads me to wonder if their time outdoors has led to these cases of allergies. But, to determine this test's accuracy and weight, I ran an independent sample t-test to pull the P-value of the data and QQ plots on those with and without allergies. The QQ plots were ran in order to test the data's distribution and normality even though the sample size is large enough.




### My Test Results
<br>


```{r T Test, message=FALSE, warning=FALSE}
pander(t.test(Outdoor_Activities_Hours ~ Allergies,
              data=HSSAllergies, 
              mu = 0, 
              alternative = "greater", 
              conf.level = 0.95),
       split.table = Inf,
       caption = "Independent Samples T-Test")

```  

Test results from an Independent Samples $t$-Test show us that the average difference between time spent outside by those with and without allergies is of little importance, $p$-value = 0.9202. With this $p$-value, we are unable to reject our null hypothesis and thus determine that those with and without allergies actually spend about the same amount of time outside. Let's see if this test is correct to use though.


```{r QQ Plot, message=FALSE, warning=FALSE}
qqPlot(Outdoor_Activities_Hours ~ Allergies,
       data=HSSAllergies)

# How do I remove the table at the bottom of the qq-plot?
# It reads ## [1] 80 75

# Why is there so much space after my Conclusion?
```

<!-- Write something about the QQ Plot not being normal, but the data size is large enough -->

Although the $p$-value gives us insufficent evidence to reject the null hypothesis, our Q-Q plots for both groups show us that the data is right-skewed. If we would like to continue using this data we need to understand that there is some abnormality within the data and cannot 100% relied upon the results. Additionally, the data of those with and without allergies is likely unrepresentative of the true population as this data was taken specifically from high school seniors. We will thus state the test to be incorrect to use. A Wilcoxon test is suggested to use for further analysis.

### Show Data {.tabset .tabset-pills .tabset-fade}

This data comes from a U.S. Census at School survey given to the High School senior students. A link to that survey can be found here:  

* https://ww2.amstat.org/censusatschool/pdfs/C@SQuestionnaire.pdf

<br>

#### Modified Data

Before I started, I ran a boxplot to see if there were any outliers within the data. I found two points at 556 and 2000 hours per week. Since there are only 168 hours in one week, I have adjusted the data to accurately show data that is equal to or less than 168 hours per week.

```{r Modified Data Table, message=FALSE, warning=FALSE}
datatable(HSSAllergies, options=list(lengthMenu = c(15,50,100)), extensions = "Responsive")
```

#### Original Data

```{r Original Data Table, message=FALSE, warning=FALSE}
datatable(HSS, options=list(lengthMenu = c(15,50,100)), extensions = "Responsive")
```




<!-- Write something about my answer, but hesitation with the normality of the data/distribution.

I decided to not do line breaks with the paragraph below. I think it looks cleaner.-->

## Conclusion
<br>
  Through an independent samples t-test, I found the $p$-value to be 0.92. Based upon the p-value, it turns out that there is insufficent evidence to reject the null hypothesis, thus showing that those with and without allergies do spend an equal amount of time outside. This stated, the results of our test cannot in good faith be relied upon because the data was taken from a very specific population and both QQ plots show that the data is right-skewed.
  An interesting follow-up study would be to first gather represenattive data and then find out if spending time outside is a factor in those who have allergies.