---
title: "Practice Final Exam"
author: "Devin Bastian"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
pacman::p_load(readr, haven, readxl, downloader, tidyverse, ggbeeswarm, mosaic, stringr, pander, DT, ggplot2, alr3, foreign, measurements, Lahman, blscrapeR, lubridate, riem, ggthemes, ggrepel, reshape2, plotly, MASS, party, lmtest)

```

```{r load_data}
# Use this R-Chunk to import all your datasets!

```

### Question 1

Question: 

Answer: 4.0341

```{r}
library(MASS)

View(cats)

?cats

summary(lm(Hwt~Bwt, data = cats))
```

### Question 2

Question: 

Answer: 

```{r}

```

### Question 3

Question: Add a regression line to this plot. Then, provide an interval that estimates what this month's Total Bill will be when it is known that this month's Gas Bill is $150.

Answer: 

```{r}
util.lm <- lm(totalbill ~ gasbill, data=Utilities)
plot(totalbill ~ gasbill, data=Utilities)
abline(util.lm)

predict(util.lm, data.frame(gasbill=150), interval="prediction")
```

### Question 4

Question: 

Answer: 

```{r}
util.lm2 <- lm(ccf ~ thermsPerDay, data=Utilities)
summary(util.lm2)
plot(ccf ~ thermsPerDay, data=Utilities)
```

### Question 5

Question: 

Answer: 

```{r}
util.lm3 <- lm(temp ~ ccf, data=Utilities)

boxCox(util.lm3)

summary(util.lm3)
```

### Question 6

Question: 

Answer: 

```{r}

CO2.loess <- loess(totalbill ~ month, data= Utilities, span=0.3, method.args=list(degree=2))
predict(CO2.loess, data.frame(month = 9), interval="prediction")
```

### Question 7

Question: 

Answer: 

```{r}

```

### Question 8

Question: 

Answer: 

```{r}
summary(lm(ccf ~ year, data = Utilities))
mean(Utilities$ccf)
```

### Question 9

Question: 

Answer: 

```{r}
pairs(airquality)
summary(lm(Wind ~ Temp, data = airquality)) 0.2045 
summary(lm(Ozone ~ Temp, data = airquality))0.4832 
summary(lm(Ozone ~ Wind, data = airquality))0.3563 
summary(lm(Solar.R ~ Wind, data = airquality))

```

### Question 10

Question: 

Answer: 

```{r}
pairs(Utilities, panel = panel.smooth, pch=16, cex=0.7, col="black")

util.lm4 <- lm(totalbill ~ gasbill, data = Utilities)

pairs(cbind(R = util.lm4$res, Fit = util.lm4$fit, Utilities), panel = panel.smooth, pch=16, cex=0.7, col="black")

```

### Question 11

Question: Use the Utilities data set to perform a regression that predicts the electric bill from the kilowatt hours of electricity usage.

Test the hypotheses that the true regression slope is $0.11. In other words, test to see if the kilowatt hours cost 11 cents per hour or not.

Answer: 

```{r}
util.lm5 <- lm(elecbill ~ kwh, data = Utilities, mu = 0.11, alternative = "two.sided", conf.level = 0.95)

summary(util.lm5)

(Estimate (b_0 or b_1) - hypothesis number)/std.Error_{b_0 or b_1}
b_0.lm <- (0.108754 - .11)/0.005816
pt(-abs(b_0.lm), 113)*2
```

### Question 12

Question: 

Answer: 

```{r}
t <- qt(.025, 115)
0.2339 + t * 0.2524
0.2339 - t * 0.2524
```

### Question 13

Question: 

Answer: 

```{r}
util.lm6 <- lm(totalbill ~ kwh + ccf + year, data = Utilities)
summary(util.lm6)
```

### Question 14

Question: Use the Utilities data set to perform a regression that shows how kilowatt hours are impacted by the year and average temperature as well as their interaction.

Within this regression there are a few minor outliers, but nothing too serious as shown by the appropriate diagnostic plots. Point 14 is currently having the most influence of any of the points on the regression, though still minimal in its influence.

Which point has the second most amount of influence on the regression?

Answer: 

```{r}
util.lm7 <- lm(kwh ~ year + temp + year:temp, data = Utilities)
summary(util.lm7)

par(mfrow=c(1,3))

plot(util.lm7, which=c(1,4,5))
```

### Question 15

Question: 

Answer: 

```{r}

```

### Question 16

Question: average billing period temperature on month of the year

Answer: 

```{r}
util.lm8 <- lm(temp ~ month + I(month^2), data = Utilities)
summary(util.lm8)
Util.lm.co <- util.lm8$coefficients

ggplot(Utilities, mapping = aes(y = temp, x = month)) +
  geom_point() +
  stat_function(fun = function(x) Util.lm.co[1] + Util.lm.co[2]*x + Util.lm.co[3]*x^2) +
  labs(title="Single Residence in Minnisota", 
       x="Month of the Year", 
       y="Monthly Gas Bill (US Dollars)")
```

### Question 17

Question: Use the Utilities data set from library(mosaic) to create a plot of a cubic regression showing the average billing period temperature on month of the year.

Select the graph from the options below that matches your plot.

Answer: 

```{r}
util.lm9 <- lm(temp ~ month + I(month^2) + I(month^3), data = Utilities)
summary(util.lm9)
Util.lm.co <- util.lm9$coefficients

ggplot(Utilities, mapping = aes(y = temp, x = month)) +
  geom_point() +
  stat_function(fun = function(x) Util.lm.co[1] + Util.lm.co[2]*x + Util.lm.co[3]*x^2 + Util.lm.co[4]*x^3) +
  labs(title="Single Residence in Minnisota", 
       x="Month of the Year", 
       y="Monthly Gas Bill (US Dollars)")
```

### Question 18

Question: 

Answer: 

```{r}
Prestige2 <- mutate(Prestige, type = as.factor(ifelse(type == "prof", "prof","other")))

summary(lm(income ~ prestige + type, data = Prestige2))



```

### Question 19

Question: 

Answer: 

```{r}

```

### Question 20

Question: 

Answer: 

```{r}
Chile.lm <- lm(income ~ population, data = Chile)
chil.co <- Chile.lm$coefficients
summary(Chile.lm)
# Check Assumptions
par(mfrow=c(1,3))
plot(Chile.lm, which=1:2)
plot(Chile.lm$residuals, main = "Residuals vs Order")

ggplot(Chile, mapping = aes(y = income, x = population)) +
  geom_point() +
  stat_function(fun = function(x) chil.co[1] + chil.co[2]*x)
```

### Question 21

Question: 

Answer: 

```{r}

summary(lm(dist ~ speed, data=cars))
```

### Question 22

Question: 

Answer: 

```{r}

```

### Question 23

Question: 

Answer: 

```{r}

```

### Question 24

Question: 

Answer: 

```{r}

predict(0.2950, 150, interval="prediction")
```

### Question 25

Question: 

Answer: 

```{r}

```