---
title: "House Prices General Advice"
author: "Me"
date: "3/18/2020"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(car)
library(pander)

train <- read.csv("~/Downloads/house-prices-advanced-regression-techniques (1)/train.csv", header=TRUE)

train <- train %>%
  mutate(TotalSqft = TotalBsmtSF + X1stFlrSF + X2ndFlrSF) %>%
  mutate(Has2ndFl = as.factor(ifelse(X2ndFlrSF > 0, "Yes", "No")))

pairs(train[,c("SalePrice", "TotalSqft", "Has2ndFl")], panel=panel.smooth)

lm1 <- lm(SalePrice ~ TotalSqft, data=train)
plot(lm1, which=1)

## Remove outliers 524 and 1299

train2 <- train[-c(524,1299), ]

lm1 <- lm(SalePrice ~ TotalSqft, data=train2)
plot(lm1, which=1)

boxCox(lm1) #suggests maybe a sqrt(sqrt(y)) 0.25...

plot(sqrt(sqrt(SalePrice)) ~ TotalSqft, data=train2, col=as.factor(Has2ndFl))


lm2 <- lm(sqrt(sqrt(SalePrice)) ~ TotalSqft, data=train2) #Not interpretable
summary(lm2)
  ## Can only interpret the effect of TotalSqft on sqrt(sqrt(SalePrice)), so you have to always talk in square root square root dollars, whatever that is!?

lm3 <- lm(log(SalePrice) ~ TotalSqft, data=train2) #Interpretable
b2 <- coef(lm3)
summary(lm3)
  ## The log is interpreted as the percentage change in Y as X increases by 1 unit.
  ## exp(slope) = %change in Y
  ## exp(b2[2]) = 1.000426 or 0.0426% increase in Y per extra Total Square Foot.
  ## exp(b2[2]*100) = 1.043537 or 4.4% increase in Y per extra 100 Total Square Feet
  ## So, if your home was 2400 SQFT and was worth $250,000, then adding on 100 more SQFT would increase the value to 1.043537*250000 = 260884.2 which is an increase of $10,884.20.


plot(SalePrice ~ TotalSqft, data=train)
b <- coef(lm2)
curve((b[1] + b[2]*x)^4, add=TRUE, col="skyblue")
curve(exp(b2[1] + b2[2]*x), add=TRUE, col="orange")

## When you transform, only the log transformation preserves interpretation.

```


## Going to 3D Space 

```{r}
pairs(cbind(R=lm3$res, Fit=lm3$fit, train2[,c("SalePrice","TotalSqft","Has2ndFl", "OverallQual","Neighborhood")]), panel=panel.smooth)

train2$Neighborhood #look at this variable
levels(train2$Neighborhood) #see names of all neighborhoods
cbind(1:25, levels(train2$Neighborhood)) #Get neighborhood numbers

## Find neighborhoods that are similar in prices
par(mai=c(1,2,.8,.1))
boxplot(SalePrice ~ Neighborhood, data=train2, horizontal=TRUE, las=1)

## Fancy sorted table of median prices
train2 %>%
  group_by(Neighborhood) %>%
  summarise(median = median(SalePrice)) %>%
  arrange(median) %>%
  pander()

train2 <- train2 %>%
  mutate(NeighborhoodGroups = case_when(Neighborhood %in% c("StoneBr","NoRidge","NridgHt") ~ "Most Expensive",
                                        Neighborhood %in% c("SawyerW","Gilbert","NWAmes","Blmngtn","CollgCr","ClearCr","Crawfor","Veenker","Somerst","Timber") ~ "Moderately Expensive",
                                        TRUE ~ "Poorer Neighborhoods"),
         NeighborhoodGroups = as.factor(NeighborhoodGroups))

pairs(cbind(R=lm3$res, Fit=lm3$fit, train2[,c("SalePrice","TotalSqft", "OverallQual","Neighborhood","NeighborhoodGroups")]), panel=panel.smooth, col=as.factor(train2$NeighborhoodGroups))


lm4 <- lm(log(SalePrice) ~ TotalSqft + NeighborhoodGroups, data=train2)
summary(lm4)

plot(SalePrice ~ TotalSqft, data=train2, col=NeighborhoodGroups)
b <- coef(lm4)
NGME=0;NGPoor=0;
curve(exp(b[1] + b[2]*x + b[3]*NGME + b[4]*NGPoor), add=TRUE, col=palette()[1])

NGME=1;NGPoor=0;
curve(exp(b[1] + b[2]*x + b[3]*NGME + b[4]*NGPoor), add=TRUE, col=palette()[2])

NGME=0;NGPoor=1;
curve(exp(b[1] + b[2]*x + b[3]*NGME + b[4]*NGPoor), add=TRUE, col=palette()[3])


## Add interactions? 
lm5 <- lm(log(SalePrice) ~ TotalSqft + NeighborhoodGroups + TotalSqft:NeighborhoodGroups, data=train2)
summary(lm5) #No need to add the interaction terms.



```



## Graphing, Interpreting, and Validating

```{r}
## Validation 
nrow(train)

## Okay to mutate, but don't filter at this stage of validation
train <- train %>%
  mutate(NeighborhoodGroups = case_when(Neighborhood %in% c("StoneBr","NoRidge","NridgHt") ~ "Most Expensive",
                                        Neighborhood %in% c("SawyerW","Gilbert","NWAmes","Blmngtn","CollgCr","ClearCr","Crawfor","Veenker","Somerst","Timber") ~ "Moderately Expensive",
                                        TRUE ~ "Poorer Neighborhoods"),
         NeighborhoodGroups = as.factor(NeighborhoodGroups))

set.seed(121)
keep <- sample(1:nrow(train), 1000)
mytrain <- train[keep,]
mytest <- train[-keep,]

## Okay to filter when running the lm...
which(rownames(mytrain) %in% c(411,633,917))
lm_train <- lm(log(SalePrice) ~ TotalSqft + LotFrontage + NeighborhoodGroups + TotalSqft:NeighborhoodGroups, data=mytrain[-c(174,264,300),])
summary(lm_train)
plot(lm_train, which=1)

## Validation

### Impute missing values
apply(mytest[,c("TotalSqft","LotFrontage","NeighborhoodGroups")], 2, function(x) sum(is.na(x)))

mytest2 <- mytest %>%
  mutate(LotFrontage = replace_na(LotFrontage, mean(LotFrontage, na.rm=TRUE)),
         BsmtQual = replace_na(BsmtQual, "Gd")) 
   #> table(mytest$BsmtQual)
   # Ex  Fa  Gd  TA 
   # 39   4 200 200 

yhat <- predict(lm_train, newdata=mytest2)

sum(is.na(yhat)) #87 missing values in yhat

Y <- log(mytest2$SalePrice) #make sure Y is on the same scale that you used in your lm

SSTO <- sum( (Y - mean(Y))^2 )
SSE <- sum( (Y - yhat)^2 )
n <- nrow(mytest2)
p <- length(coef(lm_train))
adjr2 <- 1 - (n-1)/(n-p)*SSE/SSTO #comes out to 0.651568
summary(lm_train)$adj.r.squared #original value was 0.8084779

## Best I've seen from students is 0.93 and a validation of 0.91



## Draw it
## log(SalePrice) ~ TotalSqft + LotFrontage + NeighborhoodGroups + TotalSqft:NeighborhoodGroups

plot(SalePrice ~ TotalSqft, data=mytrain)
plot(SalePrice ~ LotFrontage, data=mytrain)
plot(SalePrice ~ NeighborhoodGroups, data=mytrain) #categorical

b <- coef(lm_train)
b

plot(SalePrice ~ TotalSqft, data=mytrain, col=as.factor(NeighborhoodGroups))
LotFrontage=20;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[1])

LotFrontage=50;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[1])

LotFrontage=100;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[1])


LotFrontage=20;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[2])

LotFrontage=50;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[2])

LotFrontage=100;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[2])


LotFrontage=20;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[3], lwd=1)

LotFrontage=50;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[3], lwd=2)

LotFrontage=100;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="TotalSqft", col=palette()[3], lwd=3)



plot(SalePrice ~ LotFrontage, data=mytrain, col=as.factor(NeighborhoodGroups))

TotalSqft=1000;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[1], lwd=1)

TotalSqft=3000;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[1], lwd=2)

TotalSqft=5000;NGMostExp=0;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[1], lwd=3)


TotalSqft=1000;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[2], lwd=1)

TotalSqft=3000;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[2], lwd=2)

TotalSqft=5000;NGMostExp=1;NGPoorer=0;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[2], lwd=3)



TotalSqft=1000;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[3], lwd=1)

TotalSqft=3000;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[3], lwd=2)

TotalSqft=5000;NGMostExp=0;NGPoorer=1;
curve(exp(b[1] + b[2]*TotalSqft + b[3]*LotFrontage + b[4]*NGMostExp + b[5]*NGPoorer + b[6]*TotalSqft*NGMostExp + b[7]*TotalSqft*NGPoorer), add=TRUE, xname="LotFrontage", col=palette()[3], lwd=3)

plot(SalePrice ~ NeighborhoodGroups, data=mytrain)


## For interpretation of a log(Y) model
exp(b)



```


## Interpret it

`log(SalePrice) ~ TotalSqft + LotFrontage + NeighborhoodGroups + TotalSqft:NeighborhoodGroups`

Since everything is in log(SalePrice) space, we have to interpret each coefficient by exp(b) and it's percentage change on the SalePrice.

TotalSqft 1.000351 of the SalePrice for every 1 square foot increase. However, that isn't so interpretable, so try this instead `exp(b[2]*100)` = 1.03578 or a 3.578% increase in SalePrice for every addition 100 square feet. (Only true for the Moderately Priced Neighborhoods.)

LotFrontage 1.000902 of the SalePrice for every 1 square foot increase in LotFrontage. Or, 9.4% increase in SalePrice for each addition 100 square feet of LotFrontage

When the Neighborhood is among the "Most Expensive" types of neighborhoods, then the SalePrice is increased by `exp(b[4])` 27.8% as compared to "moderately priced" neighborhoods!

When the Neighborhood is among the "poorer" types of neighborhoods, then the SalePrice is decreased by `exp(b[5]) = 0.9096101` 9.03% as compared to "moderately priced" neighborhoods.

TotalSqft `exp((3.513394e-04-3.182852e-05)*100) = 1.032467`  increases of 100 square feet result in a 3.2% increase in the SalePrice for the Most Expensively Priced Neighborhoods.

TotalSqft `exp((3.513394e-04-6.637285e-05)*100) = 1.028907`  increases of 100 square feet result in a 2.89% increase in the SalePrice for the Most Expensively Priced Neighborhoods.






## Q & A

### Question

The column name for second floor sqr footage is "2ndFlrSF". R won't let us access it because it starts with a number. How did you deal with that?

#### Answer

```{r, eval=FALSE}
train <- train %>%
  mutate(TotalSqft = TotalBsmtSF + X1stFlrSF + X2ndFlrSF + `2ndFlrSF`)
```

### Question

How can we see the list of columns again?

```{r}
names(train)
```

### Question

Would you consider ratios to be valuable/easily interpretable (i.e. 1st floor sqft/2nd floor sqft)?

> Sometimes, these could be useful, could be dangerous. But definitely worth trying.

### Question

So, a quick question about your office hours. Will they be through zoom and only through zoom?

> They will be in my "personal meeting room" the ID: 671-623-1070

